---
title: "final-project"
author: "Yuqi Miao"
date: "12/5/2019"
output: html_document
---

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(arsenal) ## data description
library(GGally) ## data visualization
library(modelr)
library(broom)
library(viridis)
library(patchwork)
library(HH)
library(faraway)
library(leaps) # for model selection
library(caret) ## for cross validation
```

```{r}
knitr::opts_chunk$set(
  echo = T,
  message = FALSE)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

## EDA: visulization and description table

```{r}

lawsuit = read_csv("Lawsuit.csv") %>% 
    janitor::clean_names() %>% 
    mutate(
        dept = factor(dept, levels = 1:6,labels = c("Biochemistry/Molecular Biology","Physiology ","Genetics", "Pediatrics", "Medicine","Surgery")),
        gender = factor(gender, levels = c(0,1),labels = c("female","male")),
        clin = factor(clin,levels =c(1,0), labels = c("clinical","research")),
        cert = factor(cert,levels =c(1,0), labels = c("certified","not certified")),
        rank = factor(rank,levels =1:3, labels = c("Assistant", "Associate", "Full professor")),
        sal_mean = (sal94+sal95)/2
        )
    


my_controls <- tableby.control(
  total = T,
  test = T,  
  numeric.stats = c("meansd", "medianq1q3", "range"),
  stats.labels = list(
    meansd = "Mean (SD)",
    medianq1q3 = "Median (Q1, Q3)",
    range = "Min - Max"))

## create description table
tab1 <- tableby(gender~dept+clin+cert+prate+exper+rank,
                 data = lawsuit, control = my_controls, total = FALSE)

summary(tab1, title = "Descriptive statistics ", 
        text = T,  digits = 1) %>% 
  knitr::kable()   

#lawsuit %>% 
#    GGally::ggpairs(upper=list(continuous='smooth'), lower=list(continuous='cor') )

    
skimr::skim(lawsuit)   
```


## modeling

### Find confounders

```{r}
log_lawsuit = lawsuit %>% 
    mutate(ln_sal94 = log(sal94),
           ln_sal95 = log(sal95),
           ln_sal_mean = log(sal_mean)) %>% 
    dplyr::select(ln_sal94,gender,everything())

## fit the model with only gender
fit_1_sum = 
    lm(ln_sal94~gender,data = log_lawsuit) %>% 
    summary() %>%
    broom::tidy() %>% 
    mutate(model = "ln_sal94~gender")
    

## fit all other predictors with gender

p_value_list = vector("list",length = 6)
for (i in 4:9) {
  fml = as.formula(paste0("ln_sal94 ~ gender +", names(log_lawsuit)[i], collapse = ""))
  p_value_list[[i-3]] = summary(lm(fml, data = log_lawsuit))
}

## tidy the outcome into a dataframe    
p_value = vector("list",length = 6)
for (i in 1:6) {
  p_value[[i]] = broom::tidy(p_value_list[[i]]) %>% 
      mutate(
          model = paste0("ln_sal94 ~ gender +", names(log_lawsuit)[i+3], collapse = "") 
          )
}
p_value = bind_rows(fit_1_sum,p_value) 
confounder_table = p_value %>% 
    dplyr::select(model,everything()) %>% 
    filter(term == "gendermale") %>% 
    mutate(change_rate = round((estimate -0.3862374)/0.3862374,3)) %>% 
    dplyr::select(model,term,estimate,change_rate) 
    
    
    # to show the change rate of coefficient,need term, change rate, p_value
knitr::kable(confounder_table)   
```

As shown above, we choose dept, clin, cert, prate exper as our counfounder in the model as the change rate of the parameters above 10%.

## use anova to test whether rank should be included.

```{r}
fit_with_r = lm(ln_sal94 ~gender+dept+clin+cert+prate+exper+rank, data = log_lawsuit)
fit_conf = lm(ln_sal94 ~gender+dept+clin+cert+prate+exper, data = log_lawsuit)
summary(fit_conf)
anova(fit_conf,fit_with_r)
```

### rank can not be deleted.

## VIF 
```{r}
vif(fit_with_r)
```

## delete prate

### ?necessity to change reference?

### find interaction terms

* Method1: add interaction term directly into the model
```{r}
fit_int_withr = lm(ln_sal94 ~gender+dept+clin+cert+prate+exper++rank+gender*dept+gender*clin+gender*cert+gender*prate+gender*exper+gender*rank, data = log_lawsuit)
summary(fit_int_withr)
```

Thus we keep `gendermale:exper` as the interaction term

* Method2: add interaction terms saperately

Since we want to find if there is a interaction, we need to fit the interaction term in the model seperately.

```{r}
## fit one confounder with gender and its interaction; one by one

interaction_list = vector("list",length = 6)
for (i in 4:9) {
  fml = as.formula(paste0("ln_sal94 ~ gender +", names(log_lawsuit)[i],"+gender*",names(log_lawsuit)[i],collapse = ""))
  interaction_list[[i-3]] = summary(lm(fml, data = log_lawsuit))
}

## tidy the outcome into a dataframe    
interaction_p_value = vector("list",length = 6)
for (i in 1:6) {
  interaction_p_value[[i]] = broom::tidy(interaction_list[[i]]) %>% 
      mutate(
          model = paste0("ln_sal94 ~ gender +", names(log_lawsuit)[i+3], "+gender*",names(log_lawsuit)[i+3], collapse = "") 
          )
}
interaction_p_value = bind_rows(interaction_p_value) 
interaction_table = interaction_p_value %>% 
    dplyr::select(model,everything()) %>%
    filter(p.value<=0.05)
knitr::kable(interaction_table )  
```

* Make all confounders in the model
```{r}
## fit all confounders with gender and interaction one by one

interaction_list2 = vector("list",length = 6)
for (i in 4:9) {
  fml = as.formula(paste0("ln_sal94 ~ gender + dept + clin + cert + prate + exper + rank +", names(log_lawsuit)[i],"+gender*",names(log_lawsuit)[i],collapse = ""))
  interaction_list2[[i-3]] = summary(lm(fml, data = log_lawsuit))
}

## tidy the outcome into a dataframe    
interaction_p_value2 = vector("list",length = 6)
for (i in 1:6) {
  interaction_p_value2[[i]] = broom::tidy(interaction_list2[[i]]) %>% 
      mutate(
          model = paste0("ln_sal94 ~ gender + dept + clin + cert + prate + exper + rank +", names(log_lawsuit)[i+3], "+gender*",names(log_lawsuit)[i+3], collapse = "") 
          )
}
interaction_p_value2 = bind_rows(interaction_p_value2) 
interaction_table2 = interaction_p_value2 %>% 
    dplyr::select(model,everything()) %>%
    filter(p.value<=0.05)
knitr::kable(interaction_table2)  
```


### When fit the interaction term seperately, the only interaction term is gender*rank

## Do stratified analysis by rank

```{r}
rank_list = vector("list", length = 3)
for (i in 1:3) {
  rank_list[[i]] = log_lawsuit %>% 
    filter(rank == as.character(unique(log_lawsuit$rank)[i]))
}
sum_list = vector("list", length = 3)
for (i in 1:3) {
   sum_list[[i]] = summary(lm(ln_sal94 ~ gender+dept+clin+cert+prate+exper, data = rank_list[[i]]))
}

stratifid_list = map(sum_list, broom::tidy) %>%
  bind_rows() %>% 
  bind_cols(rank = rep(as.character(unique(log_lawsuit$rank)),each = 11)) %>% 
  dplyr::select(rank, everything()) %>% 
  filter(term == "gendermale")



```





```{r}
fit_8 = lm(ln_sal_mean~gender+dept+prate+exper+cert+clin+rank+gender*rank, data = log_lawsuit)
vif(fit_8)
```



## Read from articles:
1. use stepwise to find confounding:
first give an expectation of all predictors, at least direction, then use stepwise to add variable 1 by 1 to see during the process if there is a change in direction, and interprete the confounding accordingly;

2. use a saturated model including all confounders and interactions 
Interpreting with the interaction term

## change the code into functions

### find confounders


```{r}
## fit the model with only gender
fit_1_sum = 
  lm(ln_sal94~gender,data = log_lawsuit) %>% 
  summary() %>%
  broom::tidy() %>% 
  mutate(model = "ln_sal94~gender") %>% 
  dplyr::select(model,term, estimate) %>% 
  dplyr::filter(term == "gendermale")
  

## confounder function
confounder = function(y,x){
  fml =  as.formula(paste0(y," ~ gender +", x, collapse = ""))
  summary = summary(lm(fml, data = log_lawsuit))
  
  result = summary %>% 
    broom::tidy() %>% 
   mutate(
     model = paste0(y," ~ gender +", x, collapse = "")
     ) %>% 
   dplyr::select(model,term, estimate) %>% 
   dplyr::filter(term == "gendermale")
  return(result)
}

confounder_table = function(x){
  fml =  as.formula(paste0(x," ~ gender", collapse = ""))
  fit_1_sum = 
  lm(fml,data = log_lawsuit) %>% 
  summary() %>%
  broom::tidy() %>% 
  mutate(model = paste0(x,"~gender",collapse = "")) %>% 
  dplyr::select(model,term, estimate) %>% 
  dplyr::filter(term == "gendermale")
  
  first = fit_1_sum$estimate
  
  bind_rows(fit_1_sum,map2(.x = x, .y = names(log_lawsuit)[4:9], .f =confounder)) %>% ## map! 
    mutate(change = round((estimate - first)/first,3)) %>% 
    mutate(
      result = case_when(
        abs(change) == 0 ~ " ",
        abs(change) >=0.1 ~ "confounder",
        abs(change)<=0.1 ~ "not confounder"
    ))
  
}
confounder_table("ln_sal_mean")  
```

```{r}
partial_F_list =vector("list", length = 6)
fit_1 = lm(ln_sal_mean~gender, data = log_lawsuit)
fit_2 = lm(ln_sal_mean~gender+dept, data = log_lawsuit)
partial_F_list[[1]] = anova(fit_1,fit_2)
fit_3 = lm(ln_sal_mean~gender+dept+prate, data = log_lawsuit)
partial_F_list[[2]] = anova(fit_2,fit_3)
fit_4 = lm(ln_sal_mean~gender+dept+prate+exper, data = log_lawsuit)
partial_F_list[[3]] = anova(fit_3,fit_4)
fit_5 = lm(ln_sal_mean~gender+dept+prate+exper+cert, data = log_lawsuit)
partial_F_list[[4]] = anova(fit_4,fit_5)
fit_6 = lm(ln_sal_mean~gender+dept+prate+exper+cert+clin, data = log_lawsuit)
partial_F_list[[5]] = anova(fit_5,fit_6)
fit_7 = lm(ln_sal_mean~gender+dept+prate+exper+cert+clin+rank, data = log_lawsuit)
partial_F_list[[6]] = anova(fit_6,fit_7) %>% broom::tidy()
```

```{r}
partial_F_table =
  bind_rows(map(partial_F_list,broom::tidy)) %>% 
  filter(!is.na(p.value)) %>% 
  mutate(confounder = c("dept","prate","exper","cert","clin","rank")) %>% 
  dplyr::select(confounder,df,statistic,p.value)
partial_F_table
```



### find interactions

```{r}
## fit all confounders with gender and interaction one by one

interaction_list2 = vector("list",length = 6)
for (i in 4:9) {
  fml = as.formula(paste0("ln_sal94 ~ gender + dept + clin + cert + prate + exper + rank +", names(log_lawsuit)[i],"+gender*",names(log_lawsuit)[i],collapse = ""))
  interaction_list2[[i-3]] = summary(lm(fml, data = log_lawsuit))
}

## tidy the outcome into a dataframe    
interaction_p_value2 = vector("list",length = 6)
for (i in 1:6) {
  interaction_p_value2[[i]] = broom::tidy(interaction_list2[[i]]) %>% 
      mutate(
          model = paste0("ln_sal94 ~ gender + dept + clin + cert + prate + exper + rank +", names(log_lawsuit)[i+3], "+gender*",names(log_lawsuit)[i+3], collapse = "") 
          )
}
interaction_p_value2 = bind_rows(interaction_p_value2) 
interaction_table2 = interaction_p_value2 %>% 
    dplyr::select(model,everything()) %>%
    filter(p.value<=0.05)
knitr::kable(interaction_table2)  
```


```{r}
interaction_t = function(outcome, test){
  fml = as.formula(paste0(outcome,"~ gender + dept + clin + cert + prate + exper + rank +", test,"+gender*",test,collapse = ""))
  summary = summary(lm(fml,data = log_lawsuit))
  
  result = summary %>% 
    broom::tidy() %>% 
    mutate(
     add_term = paste0("gender*",test,collapse = "")
      ) %>% 
    dplyr::select(add_term,term, p.value) %>% 
    filter(str_detect(term, "gendermale:")) %>% 
    mutate(result = ifelse(p.value>0.05, "not significant", "significant"))
  
  return(result)
  
}

interaction_t_table = function(outcome){
  bind_rows(map2(.x =outcome, .y = names(log_lawsuit)[4:9], interaction_t))
}  
interaction_t_table("ln_sal94")
interaction_t_table("ln_sal95")
interaction_t_table("ln_sal_mean")
```

```{r}
interaction_F = function(outcome,test){
  fml = as.formula(paste0(outcome,"~ gender + dept + clin + cert + prate + exper + rank +", test,"+gender*",test,collapse = ""))
  anova = anova(lm(fml,data = log_lawsuit))
  
  result = anova %>% 
    broom::tidy() %>% 
    mutate(add_term = paste0("gender:",test)) %>% 
    dplyr::select(add_term, term,statistic, p.value) %>% 
    mutate(result = ifelse(p.value> 0.05,"not significant","significant")) %>% 
    filter(str_detect(term,"gender:"))
  return(result)
}

interaction_F_table = function(outcome){
  bind_rows(map2(.x = outcome, .y = names(log_lawsuit)[4:9], interaction_F))
}

interaction_F_table("ln_sal94")
interaction_F_table("ln_sal95")
interaction_F_table("ln_sal_mean")

```

## delete prate and then find interaction terms


```{r}
interaction_F = function(outcome,test){
  fml = as.formula(paste0(outcome,"~ gender + dept + clin + cert + exper + rank +", test,"+gender*",test,collapse = ""))
  anova = anova(lm(fml,data = log_lawsuit))
  
  result = anova %>% 
    broom::tidy() %>% 
    mutate(add_term = paste0("gender:",test)) %>% 
    dplyr::select(add_term, term,statistic, p.value) %>% 
    mutate(result = ifelse(p.value> 0.05,"not significant","significant")) %>% 
    filter(str_detect(term,"gender:"))
  return(result)
}

interaction_F_table = function(outcome){
  bind_rows(map2(.x = outcome, .y = names(log_lawsuit)[4:9], interaction_F))
}

interaction_F_table("ln_sal94")
interaction_F_table("ln_sal95")
interaction_F_table("ln_sal_mean")

```


```{r}
interaction_t = function(outcome, test){
  fml = as.formula(paste0(outcome,"~ gender + dept + clin + cert + exper + rank +", test,"+gender*",test,collapse = ""))
  summary = summary(lm(fml,data = log_lawsuit %>% select(-prate)))
  
  result = summary %>% 
    broom::tidy() %>% 
    mutate(
     add_term = paste0("gender*",test,collapse = "")
      ) %>% 
    dplyr::select(add_term,term, p.value) %>% 
    filter(str_detect(term, "gendermale:")) %>% 
    mutate(result = ifelse(p.value>0.05, "not significant", "significant"))
  
  return(result)
  
}

interaction_t_table = function(outcome){
  bind_rows(map2(.x =outcome, .y = names(log_lawsuit)[4:9], interaction_t))
}  
interaction_t_table("ln_sal94")
interaction_t_table("ln_sal95")
```



```{r}
log_lawsuit %>% 
  ggplot(aes(y = ln_sal94, x = exper,color = rank))+
  geom_point()+
  geom_smooth( aes(group = rank),se = F)
  
```


















