---
title: "final-project"
author: "Yuqi Miao"
date: "12/5/2019"
output: html_document
---

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(arsenal) ## data description
library(GGally) ## data visualization
library(modelr)
library(broom)
library(viridis)
library(patchwork)
library(HH)
library(faraway)
library(leaps) # for model selection
library(caret) ## for cross validation
library(car) # check the normality of variables using qqplot
library(ggcorrplot) # draw the correlation plot
```

```{r message=FALSE}
knitr::opts_chunk$set(
  echo = T,
  message = FALSE)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

## EDA: visulization and description table

```{r message=FALSE}

lawsuit = read_csv("Lawsuit.csv") %>% 
    janitor::clean_names() %>% 
    mutate(
        dept = factor(dept, levels = 1:6,labels = c("Biochemistry/Molecular Biology","Physiology ","Genetics", "Pediatrics", "Medicine","Surgery")),
        gender = factor(gender, levels = c(0,1),labels = c("female","male")),
        clin = factor(clin,levels =c(1,0), labels = c("clinical","research")),
        cert = factor(cert,levels =c(1,0), labels = c("certified","not certified")),
        rank = factor(rank,levels =1:3, labels = c("Assistant", "Associate", "Full professor")),
        sal_mean = (sal94+sal95)/2
        )
    


my_controls <- tableby.control(
  total = T,
  test = T,  
  numeric.stats = c("meansd", "medianq1q3", "range"),
  stats.labels = list(
    meansd = "Mean (SD)",
    medianq1q3 = "Median (Q1, Q3)",
    range = "Min - Max"))

## create description table
tab1 <- tableby(gender~dept+clin+cert+prate+exper+rank,
                 data = lawsuit, control = my_controls, total = FALSE)

summary(tab1, title = "Descriptive statistics ", 
        text = T,  digits = 1) %>% 
  knitr::kable()   


lawsuit %>% 
  GGally::ggpairs(
    columns = c("dept", "clin", "cert", "rank", "prate", "exper"),
    lower = list(
      continuous = "smooth",
      combo = "box",
      mapping =  aes(color = gender)
    ),
    upper = "blank") 

correlation_check = read_csv("Lawsuit.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    sal_mean = (sal94+sal95)/2
  ) %>% 
  dplyr::select(dept, gender, clin, cert, prate, exper, rank, sal_mean)

# check the correlation between the variables
cor(correlation_check) %>% 
  knitr::kable()
# print the correlation map of the variables
corr = round(cor(correlation_check), 1)
ggcorrplot(corr, method = "circle",
           hc.order = TRUE, type = "lower", lab = TRUE)

```


### Visualization - salary ~ all variables by gender 

```{r}
# dept
lawsuit %>%
  ggplot(aes(y = sal_mean, x = dept, fill = gender)) +
  geom_boxplot() + 
  labs(x = "Department", y = "Mean Salary", title = "Mean Salary versus Department by gender") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_brewer(palette="Paired")

# clin
lawsuit %>%
  ggplot(aes(y = sal_mean, x = clin, fill = gender)) +
  geom_boxplot() + 
  labs(x = "Clin", y = "Mean Salary", title = "Mean Salary versus Clin by gender") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_brewer(palette="Paired")

# cert
lawsuit %>%
  ggplot(aes(y = sal_mean, x = cert, fill = gender)) +
  geom_boxplot() + 
  labs(x = "Certified or not", y = "Mean Salary", title = "Mean Salary versus Cert by gender") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_brewer(palette="Paired")

# rank
lawsuit %>%
  ggplot(aes(y = sal_mean, x =rank, fill = gender)) +
  geom_boxplot() + 
  labs(x = "Rank", y = "Mean Salary", title = "Mean Salary versus Rank by gender") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_brewer(palette="Paired")

# prate
lawsuit %>%
  ggplot(aes(y = sal_mean, x = prate)) +
  geom_point(aes(color = gender)) + 
  labs(x = "Publication rate", y = "Mean Salary", title = "Mean Salary versus publication rate by gender") +
  scale_color_manual(values = c('#999999','#E69F00'))

# exper
lawsuit %>%
  ggplot(aes(y = sal_mean, x = exper)) +
  geom_point(aes(color = gender)) + 
  labs(x = "Number of Years since obtaining MD", y = "Mean Salary", title = "Mean Salary versus exper by gender") +
  scale_color_manual(values = c('#999999','#E69F00'))
```

## modeling

```{r}
lawsuit %>%
  ggplot(aes(x = sal_mean)) + geom_density()
#qqnorm(lawsuit$sal_mean)
#qqline(lawsuit$sal_mean)
qqPlot(lawsuit$sal_mean)
# extremely right skewed, not normally distributed
exp
full_model = lm(data = lawsuit, sal_mean ~ dept + gender + clin + cert + prate + exper + rank)
boxcox(full_model) # lamda close to 0, consider log transformation 

log_lawsuit = lawsuit %>% 
    mutate(ln_sal94 = log(sal94),
           ln_sal95 = log(sal95),
           ln_sal_mean = log(sal_mean)) %>% 
    dplyr::select(ln_sal_mean,gender,everything())

log_lawsuit %>%
  ggplot(aes(x = ln_sal_mean)) + geom_density()
#qqnorm(log_lawsuit$ln_sal_mean)
#qqline(log_lawsuit$ln_sal_mean)
qqPlot(log_lawsuit$ln_sal_mean)
# becomes nearly bell-shaped
full_model2 = lm(data = log_lawsuit, ln_sal_mean ~ dept + gender + clin + cert + prate + exper + rank)
boxcox(full_model2) # lamda close to 1
```


### Find confounders

```{r}


## fit the model with only gender
fit_1_sum = 
    lm(ln_sal_mean~gender,data = log_lawsuit) %>% 
    summary() %>%
    broom::tidy() %>% 
    mutate(model = "ln_sal_mean~gender")
    
## fit all other predictors with gender

p_value_list = vector("list",length = 6)
for (i in 4:9) {
  fml = as.formula(paste0("ln_sal_mean ~ gender +", names(log_lawsuit)[i], collapse = ""))
  p_value_list[[i-3]] = summary(lm(fml, data = log_lawsuit))
}

## tidy the outcome into a dataframe    
p_value = vector("list",length = 6)
for (i in 1:6) {
  p_value[[i]] = broom::tidy(p_value_list[[i]]) %>% 
      mutate(
          model = paste0("ln_sal_mean ~ gender +", names(log_lawsuit)[i+3], collapse = "") 
          )
}
p_value = bind_rows(fit_1_sum,p_value) 
confounder_table = p_value %>% 
    dplyr::select(model,everything()) %>% 
    filter(term == "gendermale") %>% 
    mutate(change_rate = round((estimate -0.3862374)/0.3862374,3)) %>% 
    dplyr::select(model,term,estimate,change_rate) 
    
    
    # to show the change rate of coefficient,need term, change rate, p_value
knitr::kable(confounder_table)   
```

As shown above, we choose dept, clin, cert, prate, exper as our counfounder in the model as the change rate of the parameters above 10%.

### ?necessity to change reference?

To make sure all the coefficient of the variables are significant in the model, we need to check the model with all confounders.

```{r}
fit_conf = lm(ln_sal_mean ~gender+dept+clin+cert+prate+exper, data = log_lawsuit)
summary(fit_conf) %>% 
  broom::tidy() %>% 
  filter(p.value > 0.05) %>% 
  knitr::kable()
```

We need to consider whether we need to consider whether we need to drop `prate`
Use stepwise to check the recommended model.



### find interaction terms

* Method1: add interaction term directly into the model
```{r}
fit_conf = lm(ln_sal_mean ~gender+dept+clin+cert+prate+exper+rank, data = log_lawsuit)
summary(fit_conf)
fit_int = lm(ln_sal_mean ~gender+dept+clin+cert+prate+exper+rank+gender*dept+gender*clin+gender*cert+gender*prate+gender*exper+gender*rank, data = log_lawsuit)
summary(fit_int)
```

Thus we keep `gendermale:exper` as the interaction term

* Method2: add interaction terms seperately

Since we want to find if there is a interaction, we need to fit the interaction term in the model seperately.

```{r}
## fit all other predictors with gender

interaction_list = vector("list",length = 6)
for (i in 4:9) {
  fml = as.formula(paste0("ln_sal_mean ~ gender *", names(log_lawsuit)[i], collapse = ""))
  interaction_list[[i-3]] = summary(lm(fml, data = log_lawsuit))
}

## tidy the outcome into a dataframe    
interaction_p_value = vector("list",length = 6)
for (i in 1:6) {
  interaction_p_value[[i]] = broom::tidy(interaction_list[[i]]) %>% 
      mutate(
          model = paste0("ln_sal_mean ~ gender *", names(log_lawsuit)[i+3], collapse = "") 
          )
}
interaction_p_value = bind_rows(interaction_p_value) 
interaction_table = interaction_p_value %>% 
    dplyr::select(model,everything()) %>%
    filter(p.value<=0.05)
knitr::kable(interaction_table )  


```

the model right now is

```{r}
fit_prate_rank = lm(ln_sal_mean ~ gender*rank+dept+clin+cert+exper+prate, data = log_lawsuit)
summary(fit_prate_rank)
```

stratify the rank and find out the stratified models

```{r}
# Associate
associate_fit = log_lawsuit %>% 
  filter(rank == "Associate") %>% 
  lm(ln_sal_mean ~ gender+dept+clin+cert+exper+prate, data = .)
summary(associate_fit)

# Assistant
assistant_fit = log_lawsuit %>% 
  filter(rank == "Assistant") %>% 
  lm(ln_sal_mean ~ gender+dept+clin+cert+exper+prate, data = .)
summary(assistant_fit)

# Professor
prof_fit = log_lawsuit %>% 
  filter(rank == "Full professor") %>%
  lm(ln_sal_mean ~ gender+dept+clin+cert+exper+prate, data = .)
summary(prof_fit)
```

As a result, when rank is assistant professor, gender will play a part in the amount of salary.

The preferred model now is $\hat{salary} = gender*rank+\hat{\beta_1}dept+\hat{\beta_2}clin+\hat{\beta_3}cert+\hat{\beta_4}prate+\hat{\beta_5}exper$

check whether `prate` should be kept

```{r}
fit_nprate_rank = lm(ln_sal_mean ~ gender*rank+dept+clin+cert+exper, data = log_lawsuit)
anova(fit_nprate_rank, fit_prate_rank)
```


## Heteroscedasticity, Normality

```{r}
plot(fit_nprate_rank)
plot(assistant_fit)
plot(associate_fit)
plot(prof_fit)
```

drop out the outlier in the model

```{r}
log_lawsuit_assistant = log_lawsuit %>% 
  filter(rank == "Assistant")
log_lawsuit_n68_assi = log_lawsuit_assistant[-c(68),]
fit_out68_assis = lm(ln_sal_mean ~ gender+dept+clin+cert+exper+prate, data = log_lawsuit_n68_assi)

summary(fit_out68_assis)

par(mfrow = c(2,2))
plot(fit_out68_assis)
```


## Multicollinearity

```{r}
vif(fit_prate_rank) 
vif(fit_nprate_rank)
# we should drop prate for vif>10 - serious collinearity
#our_model = 
```


## Outliers, influential points

```{r}
our_model = lm(ln_sal_mean ~gender+dept+clin+cert+exper+gender*rank, data = log_lawsuit)
summary(our_model)

# studentized residuals
stu_res = rstandard(our_model)
outlier_y = stu_res[abs(stu_res)>2.5]
outlier_y

influence.measures(our_model)

```


Annotations



## When fit the interaction term seperately, there is no interaction term.

How to distinguish?












