---
title: "final-project"
author: "Yuqi Miao"
date: "12/5/2019"
output: html_document
---

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(arsenal) ## data description
library(GGally) ## data visualization
library(modelr)
library(broom)
library(viridis)
library(patchwork)
library(HH)
library(faraway)
library(leaps) # for model selection
library(caret) ## for cross validation
```

```{r}
knitr::opts_chunk$set(
  echo = T,
  message = FALSE)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

## EDA: visulization and description table

```{r}

lawsuit = read_csv("Lawsuit.csv") %>% 
    janitor::clean_names() %>% 
    mutate(
        dept = factor(dept, levels = 1:6,labels = c("Biochemistry/Molecular Biology","Physiology ","Genetics", "Pediatrics", "Medicine","Surgery")),
        gender = factor(gender, levels = c(0,1),labels = c("female","male")),
        clin = factor(clin,levels =c(1,0), labels = c("clinical","research")),
        cert = factor(cert,levels =c(1,0), labels = c("certified","not certified")),
        rank = factor(rank,levels =1:3, labels = c("Assistant", "Associate", "Full professor")),
        sal_mean = (sal94+sal95)/2
        )
    


my_controls <- tableby.control(
  total = T,
  test = T,  
  numeric.stats = c("meansd", "medianq1q3", "range"),
  stats.labels = list(
    meansd = "Mean (SD)",
    medianq1q3 = "Median (Q1, Q3)",
    range = "Min - Max"))

## create description table
tab1 <- tableby(gender~dept+clin+cert+prate+exper+rank,
                 data = lawsuit, control = my_controls, total = FALSE)

summary(tab1, title = "Descriptive statistics ", 
        text = T,  digits = 1) %>% 
  knitr::kable()   

#lawsuit %>% 
#    GGally::ggpairs(upper=list(continuous='smooth'), lower=list(continuous='cor') )

    
   
```


## modeling

### Find confounders

```{r}
log_lawsuit = lawsuit %>% 
    mutate(ln_sal94 = log(sal94),
           ln_sal95 = log(sal95),
           ln_sal_mean = log(sal_mean)) %>% 
    dplyr::select(ln_sal94,gender,everything())

## fit the model with only gender
fit_1_sum = 
    lm(ln_sal94~gender,data = log_lawsuit) %>% 
    summary() %>%
    broom::tidy() %>% 
    mutate(model = "ln_sal94~gender")
    

## fit all other predictors with gender

p_value_list = vector("list",length = 6)
for (i in 4:9) {
  fml = as.formula(paste0("ln_sal94 ~ gender +", names(log_lawsuit)[i], collapse = ""))
  p_value_list[[i-3]] = summary(lm(fml, data = log_lawsuit))
}

## tidy the outcome into a dataframe    
p_value = vector("list",length = 6)
for (i in 1:6) {
  p_value[[i]] = broom::tidy(p_value_list[[i]]) %>% 
      mutate(
          model = paste0("ln_sal94 ~ gender +", names(log_lawsuit)[i+3], collapse = "") 
          )
}
p_value = bind_rows(fit_1_sum,p_value) 
confounder_table = p_value %>% 
    dplyr::select(model,everything()) %>% 
    filter(term == "gendermale") %>% 
    mutate(change_rate = round((estimate -0.3862374)/0.3862374,3)) %>% 
    dplyr::select(model,term,estimate,change_rate) 
    
    
    # to show the change rate of coefficient,need term, change rate, p_value
knitr::kable(confounder_table)   
```

As shown above, we choose dept, clin, cert, prate exper as our counfounder in the model as the change rate of the parameters above 10%.

## use anova to test whether rank should be included.

```{r}
fit_with_r = lm(ln_sal94 ~gender+dept+clin+cert+prate+exper+rank, data = log_lawsuit)
anova(fit_conf,fit_with_r)
```

## rank can not be deleted.

### ?necessity to change reference?

### find interaction terms

* Method1: add interaction term directly into the model
```{r}
fit_conf = lm(ln_sal94 ~gender+dept+clin+cert+prate+exper, data = log_lawsuit)
summary(fit_conf)
fit_int = lm(ln_sal94 ~gender+dept+clin+cert+prate+exper+gender*dept+gender*clin+gender*cert+gender*prate+gender*exper+gender*rank, data = log_lawsuit)
summary(fit_int)
```

Thus we keep `gendermale:exper` as the interaction term

* Method2: add interaction terms saperately

Since we want to find if there is a interaction, we need to fit the interaction term in the model seperately.

## not meaningful

```{r}
## fit all other predictors with gender

interaction_list = vector("list",length = 5)
for (i in 4:8) {
  fml = as.formula(paste0("ln_sal94 ~ gender +", names(log_lawsuit)[i],"+gender*",names(log_lawsuit)[i],collapse = ""))
  interaction_list[[i-3]] = summary(lm(fml, data = log_lawsuit))
}

## tidy the outcome into a dataframe    
interaction_p_value = vector("list",length = 5)
for (i in 1:5) {
  interaction_p_value[[i]] = broom::tidy(interaction_list[[i]]) %>% 
      mutate(
          model = paste0("ln_sal94 ~ gender +", names(log_lawsuit)[i+3], "+gender*",names(log_lawsuit)[i+3], collapse = "") 
          )
}
interaction_p_value = bind_rows(interaction_p_value) 
interaction_table = interaction_p_value %>% 
    dplyr::select(model,everything()) %>%
    filter(p.value<=0.05)
knitr::kable(interaction_table )  
```


## When fit the interaction term seperately, there is no interaction term.

How to distinguish?


```{r}
interaction_list = vector("list",length = 5)
for (i in 4:8) {
  fml = as.formula(paste0("ln_sal94 ~ gender +", names(log_lawsuit)[i],"+gender*",names(log_lawsuit)[i],collapse = ""))
  interaction_list[[i-3]] = summary(lm(fml, data = log_lawsuit))
}

## tidy the outcome into a dataframe    
interaction_p_value = vector("list",length = 5)
for (i in 1:5) {
  interaction_p_value[[i]] = broom::tidy(interaction_list[[i]]) %>% 
      mutate(
          model = paste0("ln_sal94 ~ gender +", names(log_lawsuit)[i+3], "+gender*",names(log_lawsuit)[i+3], collapse = "") 
          )
}
interaction_p_value = bind_rows(interaction_p_value) 
interaction_table = interaction_p_value %>% 
    dplyr::select(model,everything()) %>%
    filter(p.value<=0.05)
knitr::kable(interaction_table )  
```










